Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-08-03 Ben_Manuscripts/structure_paper/Draft.tex
edits for clarification
New figures and R-pi investigation

  Evidence strongly suggests that the R-spots signal is not a result of alkane chain tilt. 
  Previous literature has attributed the spots in this particular WAXS pattern as the product 
  of tilted alkane chains~\cite{feng_scalable_2014}. We measured the tilt angle of the
  alkane chains of the 280K system and showed that it equilibrates to an average tilt angle of
  -2\degree (Fig.~\ref{fig:tilt}), far from the 37\degree tilt angle previously used to 
  explain R-spots. 
  %MRS8: did you do a few different things here as sensitivity analysis to show that the finding was robust?  If so, putting some of that in would be useful.

@@

  There is not a single reason we can clearly identify for the high intensity of R-$\pi$ in our simulations.
  We used one-dimensional pair distribution functions of the centers of masses of monomer
  phenyl groups, $g(z)$, to help us understand how far apart monomers stack within monomer 
  columns and the degree to which they are correlated. These metrics helped us dissect the 
  various contributions to R-$\pi$. Figure~\ref{fig:correlation} shows $g(z)$ for all 
  systems tested.

  %MRS7: not a strong thesis sentence.
%  We plotted the one-dimensional pair distribution function, $g(z)$, for the centers
%  of mass of monomer phenyl rings (Figure~\ref{fig:correlation}). We used this 
%  information to understand how far apart monomers stack within monomer columns and
%  the degree to which they are correlated.

@@

  Monomers in the ordered basin stack closer together than those in the disordered basin
  (Table~\ref{table:correlation_length}). We calculated the equilibrated vertical distance,
  $\mathit{d}_{equil}$, between monomers using Equation~\ref{eqn:decaying_sinusoid}. 
  Monomers in the ordered basin stack ca. 0.2 and 0.3 \AA~closer than those built in the 
  disordered basin in the sandwiched and parallel displaced configurations respectively. 
  The distance between stacked monomers is greater than experiment by 0.5--0.9 \AA~across all 
  cases. This behavior is not surprising since GAFF models atoms as point charges and does not 
  appropriately model the aromatic $\pi-\pi$ interactions which would make it more energetically
  favorable for the monomers to stack closer together. Additionally, it is possible that the 
  tails prevent close stacking of monomer head groups and we do not achieve the timescales 
  necessary for them to rearrange into a necessarily more tightly packed configuration. Ordered
  basin systems are likely the best choice for transport studies.

@@

@@

  The persistent oscillatory behavior of the correlation function may in part explain the high
  intensity of R-$\pi$ relative to experiment. The way we calculate correlation length is
  unique to the rate of decay of $g(z)$ but not its exact shape. $g(z)$ for our 
  simulated systems, especially the sandwiched configurations, shows peaks which persist across
  the entire correlation function. If the peaks, at large values of z, instead decay to 1 uniformly,
  the correlation length could remain nearly the same, but R-$\pi$ would have a reduced intensity. 
  The monomer head groups in our simulations do not move enough on the timescales simulated to 
  create the noise which would cause such behavior. See Section S-\ref{S-section:rpi_ft} of the
  supporting information for illustrative examples.

>>>>

  Evidence strongly suggests that the R-spots signal is not a result of alkane chain tilt. 
  Previous literature has attributed the spots in this particular WAXS pattern as the product 
  of tilted alkane chains~\cite{feng_scalable_2014}. We measured the tilt angle of the
  alkane chains of the 280K system and showed that it equilibrates to an average tilt angle of
  -2\degree (Fig.~\ref{fig:tilt}), far from the 37\degree tilt angle previously used to 
  explain R-spots. 
  %MRS8: did you do a few different things here as sensitivity analysis to show that the finding was robust?  If so, putting some of that in would be useful.
  %BJC7: make it clear that it is the 280K. Reference all configurations where tilt angles were calculated.

@@

  % BJC8: Is it necessary to focus in on the parallel displaced system for this section? At this
  % point in the story, I haven't shown that it gives rise to R-double. I think the simple systems
  % are conceptually easier to grasp with sandwiched columns. Sliding parallel displaced columns
  % is harder to imagine, and probably not really how the columns are stacked. The parallel displacement
  % is probably less exagerated. 
  % BJC8: that said, if we go with just focusing on parallel displaced, then I can move the R-double section
  % up before this one.
  R-$\pi$ is more intense relative to the experimental system. In this section, we use the parallel displaced
  configuration in the ordered basin as an example, however all systems display this behavior.
  %since it is the most experimentally consistent system we've studied so far. 
  The intensity of R-$\pi$ in the ordered parallel displaced system is approximately 3 times
  larger than experiment (Table~\ref{table:relative_inensities_300K}).

  The full width at half maximums (FWHM) of the $q_r$ and $q_z$ cross-sections of R-$\pi$ 
  are in moderate agreement with experiment (Figure~\ref{fig:rpi_exp_comparison}). Each FWHM
  is larger than experiment. Note that the experimental data is best fit with a Lorentzian 
  function while the simulated patterns are best fit using a Gaussian function. The experimental
  pattern also has additional sources of broadening which, if deconvoluted, would increase the discrepancy with
  simulation.
  % BJC8: Can add more detail about different types of broadening - instrument / finite resolution broadening (for
  % experimental case), doppler (gaussian shaped) and pressure broadening (Lorentzian shaped)
  % See: https://www.nist.gov/pml/atomic-spectroscopy-compendium-basic-ideas-notation-data-and-formulas/atomic-spectroscopy-7

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{exp_zsection_fit.png}
  \caption{FWHM = 0.110 $\pm$ 0.003 $\AA^{-1}$}\label{fig:exp_zsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{exp_rsection_fit.png}
  \caption{FWHM = 0.792 $\pm$ 0.009 $\AA^{-1}$}\label{fig:exp_rsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{sim_zsection_fit.png}
  \caption{FWHM = 0.194 $\pm$ 0.013 $\AA^{-1}$}\label{fig:sim_zsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{sim_rsection_fit.png}
  \caption{FWHM = 0.962 $\pm$ 0.010 $\AA^{-1}$}\label{fig:sim_rsection_fit}
  \end{subfigure}
  \caption{(a) Experimentally, R-$\pi$ broadens in the $q_z$ direction with a FWHM of 0.110 $\pm$ 0.003 $\AA^{-1}$ when
  fit with a Lorentzian function. (b) R-$\pi$ broadens in the $q_r$ direction with a FWHM of 0.792 
  $\pm$ 0.009 $\AA^{-1}$ when fit with a Lorentzian function. (c) In our simulations, the FWHM of R-$\pi$ in
  the $q_z$ direction is 0.194 $\pm$ 0.013 $\AA^{-1}$ when fit with a Gaussian function, nearly twice the width
  of experiment. (d) The simulated FWHM in the $q_r$ direction is 0.962 $\pm$ 0.01 $\AA^{-1}$ when fit with
  a Gaussian function, about 20 \% larger than experiment. Note that the large spike in the center is not included
  in the fit.}\label{fig:rpi_exp_comparison}
  \end{figure}

  Here we present our understanding of what factors influence the intensity of R-$\pi$ 
  as well its FWHM in the $q_z$ and $q_r$ directions. We have identified a number of 
  potential contributions which can be modified in our model by adjusting the simulation
  conditions, forcefield or initial configuration. We assessed each using a simplified 
  model. Our simplified model represents each monomer head group as a single point and
  places them in either a parallel displaced or sandwiched configuration. We apply 
  various transformations in order to study its simulated diffraction pattern.

  As a baseline, we created a simplified model using the same amount of disorder present
  in our atomistic simulations. There are two sources of disorder to consider: thermal motion of
  atoms and quenched disorder which is created by rapid structural rearrangement during
  early equilibration and is locked into place for the remainder of the simulation time. The
  latter has important implications for the reproducibility of this work since velocity
  randomization at the beginning of equilibration will result in different quenched 
  configurations. We measured thermal disorder by measuring the standard deviation in
  the distribution of head group center of mass positions from their average positions.
  We measured quenched disorder by calculating the standard deviation of the distribution 
  of head group center of masses from their ideal positions. In the z-direction, we measured
  the deviation of the head group center of masses from an equally spaced column of head groups.
  In the xy plane, we calculated the deviation in radial position from the pore center and 
  the angular deviation from equally spaced points surrounding the pore center. The values
  measured for each system are given in Table~\ref{table:quenched_disorder}. We built a simple
  system consisting of 1000 independent configurations using these parameters. We also made
  the simple system 4 times taller in order to get better resolution in the $q_z$ direction. 
  Figure~\ref{fig:sf_qy_nocorrelation} shows the $q_y$ cross-section of R-$\pi$ generated by
  simulating the structure factor of the trajectory.

  %BJC8: Need to add thermal noise to this table
  \begin{table}[h]
  \centering
  \begin{tabular}{c|ccc}
  \toprule
  System                        & $\sigma_z$ ($\AA$) & $\sigma_\theta$ (rad) & $\sigma_r$ ($\AA$) \\
  \midrule
  Ordered Sandwiched            &        1.48        &     0.43              &     2.30           \\
  Ordered Parallel Displaced    &        1.45        &     0.43              &     2.28           \\ 
  Disordered Sandwiched         &        1.58        &     0.43              &     2.93           \\
  Disordered Parallel Displaced &        1.65        &     0.43              &     2.63           \\
  \bottomrule
  \end{tabular}
  \caption{We measured the deviation of the positions of the center of mass of head groups from
  idealized positions so that we could incorporate the same disorder into our model system.}
  \label{table:quenched_disorder} 
  \end{table}

  %BJC8: Create parallel displaced configuration: create sandwiched system. Apply uniform rotation to every other layer
  %BJC8: I'd like to make the point somewhere that the real configuration is somewhere between sandwiched and the
  % parallel displaced configuration shown here.
%  There is no major difference between R-$\pi$ of a parallel displaced system and a 
%  sandwiched system.
%  \begin{itemize}
%  	\item Intensity does not change
%  	\item 
%  \begin{itemize}

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{sf_qy_nocorrelation.png}
  \caption{}\label{fig:sf_qy_nocorrelation}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{sf_qy_correlation.png}
  \caption{}\label{fig:sf_qy_correlation.png}
  \end{subfigure} 
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{sf_qz_correlation.png}
  \caption{}\label{fig:sf_qz_correlation.png}
  \end{subfigure}   
  \caption{We generated X-ray diffraction patterns for simplified systems consisting of 
  points meant to represent monomer head groups. We added a similar amount of noise to 
  that seen in our simulations. (a) We added Gaussian noise to each particle in the z, 
  $\theta$ and r directions. The $q_y$ cross-section of R-$\pi$ contains three sharp
  peaks. (b) When we make points in each column correlated, the intensity of R-$\pi$ 
  increases but the sharp peaks remain with the same ratio of intensity between the
  center and outer peaks. (c) Correlation between scatterers in the z-direction causes
  R-$\pi$ to broaden in the $q_z$ direction.}\label{fig:sf_correlation}
  \end{figure}

  Scatterers in each column are correlated with each other. In the simplified 
  system described above, we added uncorrelated Gaussian noise in the 
  z-direction which ignores any interaction head groups might have with each
  other. To remedy this, we randomly sampled a multivariate normal distribution
  with the same variance, v, implied by Table~\ref{table:quenched_disorder}. We 
  defined the distribution using a covariance matrix such that the covariance
  in position, z, between scatterers decays exponentially from v according to the
  equation $ve^{-z/L}$, where L is the correlation length. Experimentally, the
  correlation length can be estimated as the inverse of the FWHM of the $q_z$
  cross-section of R-$\pi$.  %BJC8: needs references 
  Accordingly, the correlation length of the experimental system is $\geq$ 9 
  $\AA$. The value may be larger since the peak is likely broadened for a variety
  of other reasons. %BJC8: the reasons I can provide more detail on above.  
  %BJC8: Can show correlation effect on intensity in supporting info

  The correlation length of vertically stacked scatterers in our simulations are
  in reasonable agreement with experiment. Since it is not feasible for us to simulate
  membranes with much taller columns for increased $q_z$ resolution, and because the 
  peak shape is convoluted by the extremely intense maximum of R-$\pi$, we measured 
  correlation length by plotting the one-dimensional pair distribution function, $g(z)$,
  for the centers of mass of monomer phenyl rings (Figure~\ref{fig:correlation}). 
  %We used this 
  %information to understand how far apart monomers stack within monomer columns and
  %the degree to which they are correlated.

@@

  %BJC8: The next couple paragraphs is all old stuff that can likely be cut down a bit. But I think 
  % this is where I'd put this discussion.
  Monomers in the ordered basin stack closer together than those in the disordered basin
  (Table~\ref{table:correlation_length}). We calculated the equilibrated vertical distance,
  $\mathit{d}_{equil}$, between monomers using Equation~\ref{eqn:decaying_sinusoid}. 
  Monomers in the ordered basin stack ca. 0.2 and 0.3 \AA~closer than those built in the 
  disordered basin in the sandwiched and parallel displaced configurations respectively. 
  The distance between stacked monomers is greater than experiment by 0.5--0.9 \AA~across all 
  cases. This behavior is not surprising since GAFF models atoms as point charges and does not 
  appropriately model the aromatic $\pi-\pi$ interactions which would make it more energetically
  favorable for the monomers to stack closer together. Additionally, it is possible that the 
  tails prevent close stacking of monomer head groups and we do not achieve the timescales 
  necessary for them to rearrange into a necessarily more tightly packed configuration. Ordered
  basin systems are likely the best choice for transport studies.

@@

  Increased thermal noise in the z-direction causes a drop in the intensity of R-$\pi$, however 
  the $q_r$ profile remains unchanged. 
  \begin{itemize}
  	\item Show how I can decrease the intensity by a factor of 3-15 x by just increasing z-direction
  	thermal noise
  	\item Can put thermal noise vs. intensity in supporting info.
  	\item The peaks are still sharp though
  	\item 3 panel figure : regular intensity, 3x decrease, 15x decrease. Peak shapes all the same
  \end{itemize}

  When we allow monomer columns to move independently in the z-direction, the intensity
  of R-$\pi$ drops and the $q_r$ cross-section of the simulated diffraction patterns 
  smooths out.
  \begin{itemize}
  	\item Allowing columns to move independently smooths out the $q_r$ cross-section. Any amount
  	of dependence results in sharp peaks.
  	\item R-$\pi$ intensity decreases by up to 10x
  	\item Can put R-$\pi$ intensity vs. column independence in supporting info.
  	\item Predict level of independence in simulation (what amount of displacement cause the 
  	R-$\pi$ profile to look like that shown in Figure 8d).
  	\item Figure: (a) x \% column displacement give Figure 8d (b) Independent columns give a smooth
  	curve
  \end{itemize}

  Increasing thermal noise in the x and y directions causes R-$\pi$ to sharpen in the 
  $q_r$ direction.
  \begin{itemize}
    \item $q_r$ cross-section still too broad
    \item Adding more thermal noise sharpens it
    \item Figure: Overlay thinner FWHM plot on top of wider FWHM 
  \end{itemize}

  We have nearly the right amount of disorder / noise in our system, but correlation 
  between columns really messes up R-$\pi$.

%  \item Stacked disks (layers)
%	  \item No influence on intensity
%	  \item Some influnce on FWHM (?)

%  There is not a single reason we can clearly identify for the high intensity of R-$\pi$ in our simulations.
%  We used one-dimensional pair distribution functions of the centers of masses of monomer
%  phenyl groups, $g(z)$, to help us understand how far apart monomers stack within monomer 
%  columns and the degree to which they are correlated. These metrics helped us dissect the 
%  various contributions to R-$\pi$. Figure~\ref{fig:correlation} shows $g(z)$ for all 
%  systems tested.

  %MRS7: not a strong thesis sentence.

@@

%  The persistent oscillatory behavior of the correlation function may in part explain the high
%  intensity of R-$\pi$ relative to experiment. The way we calculate correlation length is
%  unique to the rate of decay of $g(z)$ but not its exact shape. $g(z)$ for our 
%  simulated systems, especially the sandwiched configurations, shows peaks which persist across
%  the entire correlation function. If the peaks, at large values of z, instead decay to 1 uniformly,
%  the correlation length could remain nearly the same, but R-$\pi$ would have a reduced intensity. 
%  The monomer head groups in our simulations do not move enough on the timescales simulated to 
%  create the noise which would cause such behavior. See Section S-\ref{S-section:rpi_ft} of the
%  supporting information for illustrative examples.
